Coexistence with varying environmental filtering intensity

### 2018-07-03 env_intensity.Rmd
:hammer: Reorganized script :sparkles: 3-traits case
Co-authored-by: Rekyt <matthias.grenie@ens-lyon.fr>
Co-authored-by: pierredenelle <pierre.denelle@cefe.cnrs.fr>

library(cowplot)
library(dplyr)
library(FD)
```

@@

### Moderately correlated traits
```{r}
generate_cor_traits = function(number_patches, number_species, number_other = 9,
                               cor_coef = 0.7) {

devtools::load_all()
# Initial parameters -----------------------------------------------------------
patches  <- 25   # Number of patches
species  <- 1000   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 4    # number of traits
k        <- 1.45 # Niche height
width    <- seq(10, 10, length.out = patches) # Niche breadth

@@

@@

@@

@@

```{r}
library("tidyverse")

landscape_df = function(multiple_traits_scenario, trait_types_row) {

    population_df = multiple_traits_scenario$compo[,,150] %>%
        as.data.frame() %>%
        rownames_to_column("patch") %>%  # Add patch names as column
        # Transform to tidy data.frame (define name of columns and column name for values)
        gather("species", "N", -patch)

all_landscape = lapply(names(multiple_traits), function(scenario) {
    landscape_df(multiple_traits[[scenario]], trait_types[scenario,])
})
names(all_landscape) = names(multiple_traits)

@@

# Plot showing the number of individuals at t=150 for each trait and each
# environmental value
plot_individuals_trait_environment = all_land %>%
    inner_join(traits %>%
                   as.data.frame() %>%
                   rownames_to_column("species"),
               by = "species") %>%
    gather("trait_name", "trait_value", contains("trait"), optim_env) %>%
    ggplot(aes(trait_value, patch_optim)) +
    stat_summary_2d(aes(z = N)) +
    facet_grid(trait_name ~ simulation_type,
               labeller = labeller(simulation_type =
                                       c(R2_A8 = "20% Growth\n80% Competiton",
                                         R5_A5 = "50% Growth\n50% Competiton",
                                         R8_A2 = "80% Growth\n20% Competiton"))) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Trait Value",
         y = "Optimal Trait Value",
         fill = "Number of Individuals") +
    theme_bw()
```

>>>>

# Packages ---------------------------------------------------------------------

library("cowplot")
library("tidyverse")
library("FD")

# Functions --------------------------------------------------------------------
generate_cor_traits = function(number_patches, number_species, number_other = 9,
                               cor_coef = 0.7) {

# Compute a data.frame of number of individuals in function of optimal trait
# for competition and growth
landscape_df = function(multiple_traits_scenario, trait_types_row) {

    population_df = multiple_traits_scenario$compo[,,150] %>%
        as.data.frame() %>%
        rownames_to_column("patch") %>%  # Add patch names as column
        # Transform to tidy data.frame (define name of columns and column name
        # for values)
        gather("species", "N", -patch)

@@

### Moderately correlated traits (4 traits, various scenarios)

```{r}
devtools::load_all()
# Initial parameters -----------------------------------------------------------
patches  <- 25   # Number of patches
species  <- 1000   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 4    # number of traits
k        <- 1.45 # Niche height
width    <- seq(10, 10, length.out = patches) # Niche breadth

@@

#### FD patterns

@@

#### Compute Null Models

@@

#### Number of Individuals Landscape and CWM

@@

```{r}
all_landscape = lapply(names(multiple_traits), function(scenario) {
    landscape_df(multiple_traits[[scenario]], trait_types[scenario,])
})
names(all_landscape) = names(multiple_traits)

@@

# Plot showing the number of individuals at t=150 for each trait and each
# environmental value
plot_individuals_trait_environment = all_land %>%
    inner_join(traits %>%
                   as.data.frame() %>%
                   rownames_to_column("species"),
               by = "species") %>%
    gather("trait_name", "trait_value", contains("trait"), optim_env) %>%
    ggplot(aes(trait_value, patch_optim)) +
    stat_summary_2d(aes(z = N)) +
    facet_grid(trait_name ~ simulation_type,
               labeller = labeller(simulation_type =
                                       c(R2_A8 = "20% Growth\n80% Competiton",
                                         R5_A5 = "50% Growth\n50% Competiton",
                                         R8_A2 = "80% Growth\n20% Competiton"))) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Trait Value",
         y = "Optimal Trait Value",
         fill = "Number of Individuals") +
    theme_bw()

plot_individuals_trait_environment
```

### Moderately correlated traits (3 traits)

#### Simulation Routine

```{r}
devtools::load_all()
# Initial parameters -----------------------------------------------------------
patches  <- 25   # Number of patches
species  <- 100   # Number of species
time     <- 150	 # Length of model run (generations)
initpop  <- 50   # initial population size
n_traits <- 3    # number of traits
k        <- 1.45 # Niche height
width    <- seq(10, 10, length.out = patches) # Niche breadth

# Number of trait influencing growth or competition
trait_types = c(trait1 = "R", trait2 = "RA", trait3 = "A")

# Simulations ------------------------------------------------------------------
three_traits <- multigen(traits = traits,
                         trait_type = trait_types,
                         env = env,
                         time = time,
                         species = species,
                         patches = patches,
                         composition = composition,
                         A = A, d = d,
                         k = k, width = width)

# Plot individual patches 
plot_single_patch <- directlabels::direct.label(plot_patch(three_traits$compo,
                                                           1, 150))

# Plot species richness
nb_ind_final <- three_traits$compo[,,time]
sp_rich_final <- rowSums(nb_ind_final > 2)

# sp_rich_final <- as.data.frame(sp_rich_final)
# colnames(sp_rich_final) <- names(multiple_traits)
# sp_rich_final <- tibble::rownames_to_column(sp_rich_final, "patch")
# sp_rich_final$patch <- as.numeric(gsub("patches", "", sp_rich_final$patch))
# sp_rich_final$width <- width
# 
# sp_rich_final <- tidyr::gather(sp_rich_final, "prop_type", "richness", dplyr::contains("prop"))
# 
# plot_prop_sp_rich <- ggplot(sp_rich_final, aes(1/width, richness, color = prop_type)) +
#     geom_line(size = 1) +
#     scale_color_manual(values = viridis::viridis(3)) +
#     xlab("Environmental intensity (1/width)") +
#     ylab("Patch richness")
# 
# plot_prop_sp_rich_patch <- ggplot(sp_rich_final, aes(patch, richness, color = prop_type)) +
#     geom_line(size = 1) +
#     scale_color_manual(values = viridis::viridis(3)) +
#     xlab("Patch Index") +
#     ylab("Patch richness")
```

#### Individuals Landscape and CWM

```{r}
all_land = landscape_df(three_traits, trait_types) %>%
    inner_join(traits %>%
                   as.data.frame() %>%
                   rownames_to_column("species"),
               by = "species") %>%
    group_by(patch) %>%
    mutate(rel_abund = N/sum(N)) %>%
    ungroup() %>%
    gather("trait_name", "trait_value", contains("trait"), optim_env)

# data.frame containing CWM values for all the traits, as well as growth-related
# traits (optim_env) for each patch and each simulation scenario
cwm_patch_trait = all_land %>%
    group_by(patch, trait_name) %>%
    summarise(cwm_trait = sum(rel_abund*trait_value),
              patch_optim = mean(patch_optim))

# Plot of Trait CWM in function of environmental value, colored by scenario
plot_cwm = ggplot(cwm_patch_trait, aes(patch_optim, cwm_trait)) +
    geom_abline(slope = 1, intercept = 0, linetype = 2) +
    geom_point(size = 1.5) +
    geom_smooth(method = "lm", se = FALSE) +
    facet_wrap(~trait_name, ncol = 3) +
    labs(x     = "Environmental Value",
         y     = "CWM of Trait",
         color = "Scenario") +
    theme_bw() +
    theme(legend.key.height = unit(2, "lines"))
plot_cwm

# Plot showing the number of individuals at t=150 for each trait and each
# environmental value
plot_individuals_trait_environment = all_land %>%
    ggplot(aes(trait_value, patch_optim)) +
    stat_summary_2d(aes(z = N)) +
    facet_grid(trait_name ~ .) +
    scale_fill_viridis_c(trans = "log10") +
    labs(x = "Trait Value",
         y = "Optimal Trait Value",
         fill = "Number of Individuals") +
    theme_bw()

plot_individuals_trait_environment
```
